A Lyapunov Analysis of Momentum Methods \\ in Optimization

### 2016-11-08 Draft14.tex
Update Draft14.tex

\paragraph{Problem Setting}
This paper is concerned with the class of constrained optimization problems
\begin{align}\label{eq:main-problem}
\min_{x \in \X} \; f(x),
\end{align}
where $\X \subseteq \R^d$ is a closed convex set and $f \colon \X \to \R$ is a continuously differentiable convex function. 
% We primarily focus on the case $\X = \R^d$ unless explicitly noted otherwise. 
%We also assume $f$ has a unique minimizer, $x^* \in \X$, satisfying the optimality condition $\nabla f(x^*) = 0$. 
We use the standard Euclidean norm $\|x\| = \langle x,x \rangle^{1/2}$ throughout.
We consider the general non-Euclidean setting in which the space $\X$ is endowed with a distance-generating function $h \colon \X \to \R$ that is convex and essentially smooth (i.e. $h$ is continuously differentiable in $\X$, and $\|\nabla h(x)\|_* \to \infty$ as $\|x\| \to \infty$). The function $h$ can be used to define a measure of distance in $\X$ via its Bregman divergence:
\begin{align*}
D_h(y,x) = h(y) - h(x) - \langle \nabla h(x), y-x \rangle,
\end{align*}
which is nonnegative since $h$ is convex. 
The \emph{Euclidean setting} is obtained when $h(x) = \frac{1}{2} \|x\|^2$.

@@

\paragraph{Momentum methods}
The intuition for solving problem~\eqref{eq:main-problem} with momentum methods comes from viewing optimization algorithms as discrete-time approximations to ODEs. In particular, the second-order ODE:
 \begin{align}\label{eq:heavy-ball-ode}
 m \frac{d^2x}{dt^2}& = - \nabla f(x) - m \dot \gamma_t \frac{dx}{dt}
 \end{align}
corresponds to the equations of motion of an object with mass $m$ in the presence of  friction or viscosity and a potential $f$. The trajectories of this dynamical system will tend to have directions of motion that persist over time, however for some values of $\gamma_t$, the dynamics will converge to the minimum of $f$.  In the limit as the mass goes to zero, and letting $\gamma_t \propto t/m$, we recover the gradient-flow dynamics, which is the continuous-time limit of gradient descent~\cite{Acceleration}.  Thus, the mass term here serves to \emph{accelerate} the progress towards the bottom of the well.

@@

Until recently, this ODE perspective on optimization algorithms was largely abandoned in favor of Nesterov's estimate sequence framework.  However, there has been a recent resurgence in interest in ODE-based analyses.  Su, Boyd and Cand\`es~\cite{SuBoydCandes14}, showed that the continuous-time limit of Nesterov's accelerated gradient descent method corresponds to the following second-order ODE, 
 \begin{equation}\label{Eq:SuBoydCandes}
 \ddot X_t + \frac{r}{t} \dot X_t + \nabla f(X_t) = 0,\,\,\,\,r \geq 3.
 \end{equation}
This is equivalent to~\eqref{eq:heavy-ball-ode}, where the dynamics model an object with unit mass and the damping term is logarithmic, $\gamma_t = r\log t$.  They also established a convergence rate of $O(1/t^2)$ for this continuous-time system via a Lyapunov function argument. This matches the $O(1/\epsilon k^2)$ convergence rate of accelerated gradient descent, given the discretization $t= \sqrt\epsilon k$. Thus, the convergence rate is maintained in passing from continuous time to discrete time. Two things remain unclear from their analysis however:  they introduce an additional sequence $\{y_k\}_{k=1}^\infty$ while discretizing the ODE~\cite[Sec 2]{Acceleration},  which does not correspond to a straightforward discretization technique; and they do not demonstrate what connection---if any---there is between the Lypaunov argument used to analyze the continuous-time dynamics and the technique of estimate sequences used to analyze the discrete-time algorithm.

@@

Wibisono, Wilson, and Jordan~\cite{Acceleration} introduced a class of dynamical systems whose discretizations give rise to a general family of accelerated gradient algorithms, including accelerated gradient and mirror descent. These algorithms are obtained as a discretization of the following second-order ODE:
 \begin{align} \label{Eq:ELBreg} 
\frac{d}{dt} \nabla h(X_t + e^{-\alpha_t} \dot X_t) = -e^{\alpha_t+\beta_t} \nabla f(X_t)\,,
\end{align}
where $h$ is a strongly convex function associated with the geometry of the space and where $\alpha_t$ and $\beta_t$ are smooth functions of time.  In the case where $\alpha_t=-\log m$, and $\beta_t = \log m$,  the continuous-time limit $m \rightarrow 0$ of this ODE recovers the mirror descent dynamics~\cite[(78)]{Acceleration}, with $h$ playing the role of the distance-generating function.  Indeed, the original motivation for mirror descent by Nemirovski and Yudin was from this continuous perspective, and they demonstrated a Lyapunov argument for the convergence of the continuous-time dynamics (see Section 3.1 of Nemirovski and Yudin~\cite{NemirovskiiYudin}).

@@

In this paper, we are interested in obtaining Lyapunov functions for dynamical systems designed for optimization.  In optimization, we are typically interested in the~\emph{rate of convergence} of a method, not mere convergence.  To facilitate such rate analysis, we study time-dependent Lyapunov functions.  Though it is unconventional to have Lyapunov functions depend explicitly on time, like the dynamics~\eqref{Eq:ELBreg}, our Lyapunov function is parameterized explicitly by $\alpha_t$, $\beta_t$, and $\gamma_t$.  The family of dynamics that obey the ideal scaling constraints~\eqref{Eq:IdeSca} is invariant under time reparemeterizations.  Such time-invariance properties provide more flexibility in algorithm design and analysis, as we will see throughout this paper.

@@

\paragraph{General Accelerated Mirror Descent} In Wibisono {\em et al.}~\cite{Acceleration}, the following Lyapunov function 
\begin{align}\label{Eq:E}
\E_t \,=\, e^{\beta_t} (f(X_t)-f(x^*)) + D_h\left(x^*, \, X_t + e^{-\alpha_t} \dot X_t\right),
\end{align}
was introduced to analyze the flow that forms the basis for accelerated gradient descent and its non-Euclidean variants.
%  Notice, assuming we start from rest $\dot X_0 = 0$, the  Lyapunov property allows us to conclude,
 %\begin{equation*}
 %f(X_t) - f(x^\ast)\leq \frac{e^{\beta_0} (f(X_0) - f(x^\ast) )+ D_h(x^\ast, X_0)}{e^{\beta_t}}.
 %\end{equation*}
In Section~\ref{Sec:MomLyap}, we show how this Lyapunov function emerges from a general framework for analyzing momentum dynamics; furthermore, we show how to discretize the Lyapunov function and thereby obtain a tool which can be used to analyze a variety of algorithms that are discretizations of momentum dynamics.
%MJ:  I don't view this footnote as necessary; that derivation was in supplementary information \footnote{The same derivation of the Lyapunov function was added to~\cite[App. C]{Acceleration}}
%
%
%
%ACCC GRAD STRONG
\paragraph{Accelerated Gradient Descent (Strong Convexity)}
In this paper, we show that the following function 
   \begin{equation}\label{eq:general-sc-lyap}
  \E_t =  e^{\beta_t} \left(f(X_t) - f(x^\ast) + \mu D_h\Big(x^\ast ,X_t + e^{-\alpha_t} \dot X_t\Big)\right),
 \end{equation}
is a Lyapunov function for a family of momentum dynamics when $f$ is $\mu$-strongly convex ($D_f(y,x) \geq \mu D_h(y,x)$) and $e^{\alpha_t} = \dot \beta_t$.  In Section~\ref{Sec:StrongConv}, we derive this Lyapunov function and show how it can be used to demonstrate a linear rate of convergence for the family of accelerated gradient descent dynamics.  (For simplicity, we focus on the setting in which $h = \frac{1}{2}\|\cdot\|^2$ is Euclidean.) We also show that a particular discretization of the family of dynamics provides an algorithm for which a commensurate discretization of~\eqref{eq:general-sc-lyap} is a Lyapunov function.

@@

Note that in both the strongly convex and weakly convex setting, the structure of the Lyapunov function allows us to derive a rate of convergence for the optimality gap $f(X_t) - f(x^\ast)$. In particular, the Lyapunov property $\E_t \leq \E_0$ implies 
\begin{equation}\label{eq:Guarantee1}
f(X_t) - f(x^\ast) \leq \frac{e^{\beta_0}(f(X_0) - f(x^\ast)) + D_h(x^\ast, X_0 + e^{-\alpha_0} \dot X_0)}{e^{\beta_t}}
\end{equation}
and 
\begin{equation}
f(X_t) - f(x^\ast) \leq \frac{e^{\beta_0}(f(X_0) - f(x^\ast) +\mu D_h(x^\ast ,X_0 -e^{-\alpha_0} \dot X_0))}{e^{\beta_t}},
\end{equation}
for~\eqref{Eq:E} and \eqref{eq:general-sc-lyap} respectively. The typical convention is that the dynamics start from rest $\dot X_0 = 0$. Notice that we obtain the same rates of convergence in these two settings. However, the family of dynamics corresponding to the Lyapunov function in the strongly convex setting uses the strong convexity of $f$, and this allows us to obtain a tighter bound in discrete time.

\section{Momentum Methods}
\label{Sec:MomLyap}
In this section, we introduce two families of momentum dynamics, both which are generated by functionals defined on continuous-time curves. We illustrate how to derive a Lyapunov function from these dynamics. Next, we show how these dynamics can be discretized to obtain optimization algorithms. The same discrete-time mappings can be applied to the Lyapunov function, resulting in a tool which can be used to analyze the discrete-time algorithms. We provide examples of this procedure for the quasi-monotone subgradient algorithm, accelerated gradient descent, the conditional gradient algorithm, and several other novel algorithms. 
 \subsection{Lyapunov Analysis}
We study dynamics that are based on the Euler-Lagrange equation \eqref{Eq:ELBreg} in the setting where the ideal scaling~\eqref{Eq:IdeScaBet} holds with equality~$\dot \beta_t = e^{\alpha_t}$. Notice, now the entire Bregman Lagrangian is parameterized by $\beta_t$, which determines the convergence rate:
 \begin{subequations}\label{Eq:EL}
 \begin{align}
 Z_t &= X_t + \frac{1}{\dot \beta_t} \dot X_t \label{Eq:ELZ},\\
 \frac{d}{dt}\nabla h(Z_t) &= -\dot \beta_t e^{\beta_t} \nabla f(X_t)\label{Eq:ELH}.
 \end{align}
 \end{subequations}
We begin by demonstrating how to derive the Lyapunov function~\eqref{Eq:E} for the momentum dynamics~\eqref{Eq:EL}; this derivation is similar in spirit to the Lyapunov analysis of mirror descent by Nemirovski and Yudin. We first point the following nice identity, 
\begin{subequations}
\begin{align}
 \frac{d}{dt}D_h\left(x, Z_t\right)  &= \frac{d}{dt}\left( h(x) - h(Z_t) - \langle \nabla h(Z_t), x- Z_t\rangle\right)  \notag\\
& = - \langle \nabla h(Z_t), \dot Z_t\rangle -\left\langle \frac{d}{dt} \nabla h(Z_t),x - Z_t\right\rangle  + \langle \nabla h(Z_t), \dot Z_t\rangle \notag\\ 
 &= -\left\langle \frac{d}{dt} \nabla h\left(Z_t\right), x - Z_t\right\rangle. \, \notag%\label{eq:apply-identity}\\
 \end{align*}
 \end{subequations}
  Using this identity, we have the following argument:
%\begin{small}
\begin{subequations}\label{Eq:LyapAnal}
\begin{align}
 \frac{d}{dt}D_h\left(x, Z_t\right)  %&= \frac{d}{dt}\left( h(x) - h(Z_t) - \langle \nabla h(Z_t), x- Z_t\rangle\right)  \notag\\
%& = - \langle \nabla h(Z_t), \dot Z_t\rangle -\left\langle \frac{d}{dt} \nabla h(Z_t),x - Z_t\right\rangle  + \langle \nabla h(Z_t), \dot Z_t\rangle \notag\\ 
 &= -\left\langle \frac{d}{dt} \nabla h\left(Z_t\right), x - Z_t\right\rangle \, \notag\\%\label{eq:apply-identity}\\
&=   \dot \beta_t e^{\beta_t}\left\langle\nabla f(X_t), x  - X_{t} - \frac{1}{\dot \beta_t} \dot X_{s} \right\rangle  \, \label{eq:apply-dynamics}   \\
& =   \dot \beta_t e^{\beta_t} \langle \nabla f(X_t), x - X_t \rangle\, - e^{ \beta_t} \langle \nabla f(X_t), \dot X_t\rangle dt\, \notag   \\
& = \dot \beta_t e^{\beta_t} \langle \nabla f(X_t), x - X_t \rangle\,  -   \frac{d}{dt}\left(e^{ \beta_t} f(X_t)\right)\,   + \dot \beta_t e^{\beta_t}  f(X_t)  \notag\\
&= \dot \beta_t e^{\beta_t} [f(X_t)+ \langle \nabla f(X_t), x - X_t \rangle] -\frac{d}{dt}\left(e^{ \beta_t} f(X_t)\right)\,  \notag  \\ %\label{Eq:Line9}\\
& \leq  \dot \beta_t e^{\beta_t}f(x) - \frac{d}{dt}\left(e^{ \beta_t} f(X_t)\right)\, \label{eq:apply-convexity}\\
& = - \frac{d}{dt} \left(e^{\beta_t}\left(f(X_t) - f(x)\right)\right) \label{Eq:cond1}.
\end{align}
\end{subequations}
%\end{small}
Here~\eqref{eq:apply-dynamics} uses the momentum dynamics%~\eqref{Eq:EL}.
~\eqref{Eq:ELH} and~\eqref{Eq:ELZ}. 
The inequality~\eqref{eq:apply-convexity} follows from the convexity of $f$. Rearranging terms and taking $x = x^\ast$, we have shown that the function~\eqref{Eq:E} has nonpositive derivative for all $t$ and is hence a Lyapunov function for the family of momentum dynamics~\eqref{Eq:EL}.

 The structure of this derivation provides a tool for us to analyze several discretizations of momentum dynamics~\eqref{Eq:ELBreg}.

@@

 In what follows, we analyze three combinations of the implicit and explicit Euler methods using our Lyapunov framework. Furthermore, we show there are at least two slightly different methods, which do not comport to a straight-forward discretization technique, that can also be analyzed using the above Lyapunov analysis. 

@@

 We give a short analysis of the result of implicit Euler discretization of both~\eqref{Eq:Z} and~\eqref{Eq:X}. We can write the algorithm as follows,

@@

\noindent To analyze this algorithm, we make the same continuous to discrete time identifications to~\eqref{Eq:E}, which results in the following 
discrete-time energy function,

@@

 Euler discretization~\eqref{Eq:ImplicEul}. Using the discrete Lyapunov function~\eqref{Eq:DiscLyapFunc1}, the following bound holds,

@@

&\quad  + A_k (f(x_k) - f(x_{k+1}))- D_h(z_{k+1}, z_k),\notag \\
&=- D_h(z_{k+1}, z_k) \notag

@@

\noindent Taking $x = x^\ast$, writing $E_k \leq E_0$ allows us to obtain a convergence rate guarantee analogous to~\eqref{eq:Guarantee1},

@@

f(x_k) - f(x^\ast) \leq \frac{D_h(x^\ast, x_0) + A_0 (f(x_0) - f(x^\ast))}{A_k},

@@

involves applying the implicit Euler method to the mirror descent dynamic, and both yield a $O(1/A_k)$ convergence rate. 

>>>>

\paragraph{Problem Setting}
This paper is concerned with the class of constrained optimization problems
\begin{align}\label{eq:main-problem}
\min_{x \in \X} \; f(x),
\end{align}
where $\X \subseteq \R^d$ is a closed convex set and $f \colon \X \to \R$ is a continuously differentiable convex function. 
% We primarily focus on the case $\X = \R^d$ unless explicitly noted otherwise. 
%We also assume $f$ has a unique minimizer, $x^* \in \X$, satisfying the optimality condition $\nabla f(x^*) = 0$. 
We use the standard Euclidean norm $\|x\| = \langle x,x \rangle^{1/2}$ throughout.
We consider the general non-Euclidean setting in which the space $\X$ is endowed with a distance-generating function $h \colon \X \to \R$ that is convex and essentially smooth (i.e., $h$ is continuously differentiable in $\X$, and $\|\nabla h(x)\|_* \to \infty$ as $\|x\| \to \infty$). The function $h$ can be used to define a measure of distance in $\X$ via its Bregman divergence:
\begin{align*}
D_h(y,x) = h(y) - h(x) - \langle \nabla h(x), y-x \rangle,
\end{align*}
which is nonnegative since $h$ is convex. 
The \emph{Euclidean setting} is obtained when $h(x) = \frac{1}{2} \|x\|^2$.

@@

\paragraph{Momentum methods}
The intuition for solving problem~\eqref{eq:main-problem} with momentum methods comes from viewing optimization algorithms as discrete-time approximations to ODEs. In particular, the second-order ODE:
 \begin{align}\label{eq:heavy-ball-ode}
 m \frac{d^2x}{dt^2}& = - \nabla f(x) - m \dot \gamma_t \frac{dx}{dt}
 \end{align}
corresponds to the equations of motion of an object with mass $m$ in the presence of  friction or viscosity, captured by the derivative of a function $\gamma_t$, and a potential $f$. The trajectories of this dynamical system will tend to have directions of motion that persist over time; however, for some values of $\gamma_t$, the dynamics will converge to the minimum of $f$.  In the limit as the mass goes to zero, and letting $\gamma_t \propto t/m$, we recover the gradient-flow dynamics, which is the continuous-time limit of gradient descent~\cite{Acceleration}.  Thus, the mass term here serves to \emph{accelerate} the progress towards the bottom of the well.

@@

Until recently, this ODE perspective on optimization algorithms was largely abandoned in favor of Nesterov's estimate sequence framework.  However, there has been a recent resurgence in interest in ODE-based analyses.  Su, Boyd and Cand\`es~\cite{SuBoydCandes14}, showed that the continuous-time limit of Nesterov's accelerated gradient descent method corresponds to the following second-order ODE, 
 \begin{equation}\label{Eq:SuBoydCandes}
 \ddot X_t + \frac{r}{t} \dot X_t + \nabla f(X_t) = 0,\,\,\,\,r \geq 3.
 \end{equation}
This is equivalent to~\eqref{eq:heavy-ball-ode}, where the object being modeled has unit mass and the damping term is logarithmic, $\gamma_t = r\log t$.  They also established a convergence rate of $O(1/t^2)$ for this continuous-time system via a Lyapunov function argument. This matches the $O(1/\epsilon k^2)$ convergence rate of accelerated gradient descent, given the discretization $t= \sqrt\epsilon k$. Thus, the convergence rate is maintained in passing from continuous time to discrete time. Two things remain unclear from their analysis however:  they introduce an additional sequence $\{y_k\}_{k=1}^\infty$ while discretizing the ODE~\cite[Sec 2]{Acceleration},  which does not correspond to a straightforward discretization technique; and they do not demonstrate what connection---if any---there is between the Lyapunov argument used to analyze the continuous-time dynamics and the technique of estimate sequences used to analyze the discrete-time algorithm.

@@

Wibisono, Wilson, and Jordan~\cite{Acceleration} introduced a class of dynamical systems whose discretizations give rise to a general family of accelerated gradient algorithms, including accelerated gradient and mirror descent. These algorithms are obtained as a discretization of the following second-order ODE:
 \begin{align} \label{Eq:ELBreg} 
\frac{d}{dt} \nabla h(X_t + e^{-\alpha_t} \dot X_t) = -e^{\alpha_t+\beta_t} \nabla f(X_t)\,,
\end{align}
where $h$ is a strongly convex function associated with the geometry of the space and where $\alpha_t$ and $\beta_t$ are smooth functions of time.  In the case where $\alpha_t=-\log m$, and $\beta_t = \log m$,  the continuous-time limit $m \rightarrow 0$ of this ODE recovers the mirror descent dynamics~\cite{Acceleration}, with $h$ playing the role of the distance-generating function.  Indeed, the original motivation for mirror descent by Nemirovski and Yudin was from this continuous perspective, and they demonstrated a Lyapunov argument for the convergence of the continuous-time dynamics (see Section 3.1 of Nemirovski and Yudin~\cite{NemirovskiiYudin}).

@@

In this paper, we are interested in obtaining Lyapunov functions for dynamical systems designed for optimization.  In optimization, we are typically interested in the~\emph{rate of convergence} of a method, not mere convergence.  To facilitate such rate analysis, we study time-dependent Lyapunov functions.  Though it is unconventional to have Lyapunov functions depend explicitly on time, our Lyapunov function is parameterized explicitly by $\alpha_t$, $\beta_t$, and $\gamma_t$, reflecting the dependence of the dynamics~\eqref{Eq:ELBreg} on these functions.

The family of dynamics that obey the ideal scaling constraints~\eqref{Eq:IdeSca} is invariant under time reparametrizations.  Such time-invariance properties provide more flexibility in algorithm design and analysis, as we will see throughout this paper.

@@

\paragraph{General Accelerated Mirror Descent} In Wibisono {\em et al.}~\cite{Acceleration}, the following Lyapunov function:
\begin{align}\label{Eq:E}
\E_t \,=\, e^{\beta_t} (f(X_t)-f(x^*)) + D_h\left(x^*, \, X_t + e^{-\alpha_t} \dot X_t\right),
\end{align}
was introduced to analyze the flow that forms the basis for accelerated mirror descent.
%  Notice, assuming we start from rest $\dot X_0 = 0$, the  Lyapunov property allows us to conclude,
 %\begin{equation*}
 %f(X_t) - f(x^\ast)\leq \frac{e^{\beta_0} (f(X_0) - f(x^\ast) )+ D_h(x^\ast, X_0)}{e^{\beta_t}}.
 %\end{equation*}
In Section~\ref{Sec:MomLyap}, we show how this Lyapunov function emerges from a general framework for analyzing momentum dynamics; furthermore, we show how to discretize the Lyapunov function and thereby obtain a tool which can be used to analyze a variety of algorithms that are discretizations of momentum dynamics.
%MJ:  I don't view this footnote as necessary; that derivation was in supplementary information \footnote{The same derivation of the Lyapunov function was added to~\cite[App. C]{Acceleration}}
%
%
%
%ACCC GRAD STRONG
\paragraph{Accelerated Gradient Descent (Strong Convexity)}
In this paper, we show that the following function 
   \begin{equation}\label{eq:general-sc-lyap}
  \E_t =  e^{\beta_t} \left(f(X_t) - f(x^\ast) + \mu D_h\Big(x^\ast ,X_t + e^{-\alpha_t} \dot X_t\Big)\right),
 \end{equation}
is a Lyapunov function for a family of momentum dynamics when $f$ is $\mu$-strongly convex ($D_f(y,x) \geq \mu D_h(y,x)$) and $e^{\alpha_t} = \dot \beta_t$.  In Section~\ref{Sec:StrongConv}, we derive this Lyapunov function and show how it can be used to demonstrate a linear rate of convergence for the family of accelerated gradient descent dynamics.  (For simplicity, we focus on the setting in which $h = \frac{1}{2}\|\cdot\|^2$ is Euclidean.) We also show that a particular discretization of the family of dynamics provides an algorithm for which a commensurate discretization of~\eqref{eq:general-sc-lyap} is a Lyapunov function.

@@

Note that in both the strongly convex and weakly convex setting, the structure of the Lyapunov function allows us to derive a rate of convergence for the optimality gap $f(X_t) - f(x^\ast)$. In particular, the Lyapunov property $\E_t \leq \E_0$ implies 
\begin{equation}\label{eq:Guarantee1}
f(X_t) - f(x^\ast) \leq \frac{e^{\beta_0}(f(X_0) - f(x^\ast)) + D_h(x^\ast, X_0 + e^{-\alpha_0} \dot X_0)}{e^{\beta_t}},
\end{equation}
and 
\begin{equation}
f(X_t) - f(x^\ast) \leq \frac{e^{\beta_0}(f(X_0) - f(x^\ast) +\mu D_h(x^\ast ,X_0 -e^{-\alpha_0} \dot X_0))}{e^{\beta_t}},
\end{equation}
for~\eqref{Eq:E} and \eqref{eq:general-sc-lyap} respectively. The typical convention is that the dynamics start from rest $\dot X_0 = 0$. Notice that we obtain the same rates of convergence in these two settings. However, the family of dynamics corresponding to the Lyapunov function in the strongly convex setting exploits the strong convexity of $f$, and this allows us to obtain a tighter bound in discrete time.

\section{Momentum Methods}
\label{Sec:MomLyap}
In this section, we introduce two families of momentum dynamics, both which are generated by functionals defined on continuous-time curves. We illustrate how to derive a Lyapunov function from these dynamics. Next, we show how these dynamics can be discretized to obtain optimization algorithms. The same discrete-time mappings can be applied to the Lyapunov function, resulting in a tool which can be used to analyze the discrete-time algorithms. We provide examples of this procedure for the quasi-monotone subgradient algorithm, accelerated gradient descent, the conditional gradient algorithm, and several other novel algorithms. 
 \subsection{Lyapunov Analysis}
We study dynamics that are based on the Euler-Lagrange equation \eqref{Eq:ELBreg} in the setting where the ideal scaling~\eqref{Eq:IdeScaBet} holds with equality~$\dot \beta_t = e^{\alpha_t}$. This implies that the Bregman Lagrangian is parameterized by $\beta_t$, which determines the convergence rate:
 \begin{subequations}\label{Eq:EL}
 \begin{align}
 Z_t &= X_t + \frac{1}{\dot \beta_t} \dot X_t \label{Eq:ELZ},\\
 \frac{d}{dt}\nabla h(Z_t) &= -\dot \beta_t e^{\beta_t} \nabla f(X_t)\label{Eq:ELH}.
 \end{align}
 \end{subequations}
We begin by demonstrating how to derive the Lyapunov function~\eqref{Eq:E} for the momentum dynamics~\eqref{Eq:EL}; this derivation is similar in spirit to the Lyapunov analysis of mirror descent by Nemirovski and Yudin.  We have:
\begin{subequations}
\begin{align}
 \frac{d}{dt}D_h\left(x, Z_t\right)  &= \frac{d}{dt}\left( h(x) - h(Z_t) - \langle \nabla h(Z_t), x- Z_t\rangle\right)  \notag\\
& = - \langle \nabla h(Z_t), \dot Z_t\rangle -\left\langle \frac{d}{dt} \nabla h(Z_t),x - Z_t\right\rangle  + \langle \nabla h(Z_t), \dot Z_t\rangle \notag\\ 
 &= -\left\langle \frac{d}{dt} \nabla h\left(Z_t\right), x - Z_t\right\rangle. \, \notag%\label{eq:apply-identity}\\
 \end{align*}
 \end{subequations}
Using this identity, we obtain the following argument:
%\begin{small}
\begin{subequations}\label{Eq:LyapAnal}
\begin{align}
 \frac{d}{dt}D_h\left(x, Z_t\right)  %&= \frac{d}{dt}\left( h(x) - h(Z_t) - \langle \nabla h(Z_t), x- Z_t\rangle\right)  \notag\\
%& = - \langle \nabla h(Z_t), \dot Z_t\rangle -\left\langle \frac{d}{dt} \nabla h(Z_t),x - Z_t\right\rangle  + \langle \nabla h(Z_t), \dot Z_t\rangle \notag\\ 
 &= -\left\langle \frac{d}{dt} \nabla h\left(Z_t\right), x - Z_t\right\rangle \, \notag\\%\label{eq:apply-identity}\\
&=   \dot \beta_t e^{\beta_t}\left\langle\nabla f(X_t), x  - X_{t} - \frac{1}{\dot \beta_t} \dot X_{s} \right\rangle  \, \label{eq:apply-dynamics}   \\
& =   \dot \beta_t e^{\beta_t} \langle \nabla f(X_t), x - X_t \rangle\, - e^{ \beta_t} \langle \nabla f(X_t), \dot X_t\rangle dt\, \notag   \\
& = \dot \beta_t e^{\beta_t} \langle \nabla f(X_t), x - X_t \rangle\,  -   \frac{d}{dt}\left(e^{ \beta_t} f(X_t)\right)\,   + \dot \beta_t e^{\beta_t}  f(X_t)  \notag\\
&= \dot \beta_t e^{\beta_t} [f(X_t)+ \langle \nabla f(X_t), x - X_t \rangle] -\frac{d}{dt}\left(e^{ \beta_t} f(X_t)\right)\,  \notag  \\ %\label{Eq:Line9}\\
& \leq  \dot \beta_t e^{\beta_t}f(x) - \frac{d}{dt}\left(e^{ \beta_t} f(X_t)\right)\, \label{eq:apply-convexity}\\
& = - \frac{d}{dt} \left(e^{\beta_t}\left(f(X_t) - f(x)\right)\right) \label{Eq:cond1}.
\end{align}
\end{subequations}
%\end{small}
Here~\eqref{eq:apply-dynamics} uses the momentum dynamics%~\eqref{Eq:EL}.
~\eqref{Eq:ELH} and~\eqref{Eq:ELZ}. 
The inequality~\eqref{eq:apply-convexity} follows from the convexity of $f$. Rearranging terms and taking $x = x^\ast$, we have shown that the function~\eqref{Eq:E} has nonpositive derivative for all $t$ and is hence a Lyapunov function for the family of momentum dynamics~\eqref{Eq:EL}.

 The structure of this derivation provides a tool for the analysis of several discretizations of momentum dynamics~\eqref{Eq:ELBreg}.

@@

 In the following, we analyze three combinations of the implicit and explicit Euler methods using our Lyapunov framework. Furthermore, we show there are at least two slightly different methods, which do not comport to a straightforward discretization technique, that can also be analyzed using our Lyapunov analysis. 

@@

 We give a short analysis of the result of implicit Euler discretization of both~\eqref{Eq:Z} and~\eqref{Eq:X}. We can write the algorithm as follows:

@@

\noindent To analyze this algorithm, we make the same continuous to discrete time identifications as in~\eqref{Eq:E}, which results in the following discrete-time energy function:

@@

 Euler discretization~\eqref{Eq:ImplicEul}. Using the discrete Lyapunov function~\eqref{Eq:DiscLyapFunc1}, the following bound holds:

@@

&\quad  + A_k (f(x_k) - f(x_{k+1}))- D_h(z_{k+1}, z_k)\notag \\
&=- D_h(z_{k+1}, z_k), \notag

@@

\noindent Taking $x = x^\ast$, and writing $E_k \leq E_0$ allows us to obtain a convergence rate guarantee analogous to~\eqref{eq:Guarantee1}:

@@

f(x_k) - f(x^\ast) \leq \frac{D_h(x^\ast, x_0) + A_0 (f(x_0) - f(x^\ast))}{A_k}.

@@

involves applying the implicit Euler method to the mirror descent dynamics, and both yield a $O(1/A_k)$ convergence rate. 
